### 2012-02-12 paper.tex
Small corrections throughout the manuscript. The discussion is now at least decently self-consistent.

CO oxidation is one of the most studied reactions due to its importance in both automotive catalysis and in the removal of trace CO in hydrogen streams. A large effort and a large number of experiments with a variety of surface sensitive techniques have been conducted on single crystal surfaces greatly improving the understanding of this reaction. Single crystals surfaces, however, can be very much different from industrial catalysts where supported nanoparticles are used to both increase surface area hence bringing material costs down and to increase the reactivity of the material by increasing the number of active (under-coordinated?) sites.

Here we report oscillatory behavior of CO oxidation on size-selected platinum nanoparticles supported on a silicon dioxide substrate at atmospheric pressure. Oscilating reactions on Platiunum at UHV-conditions has been studied extensively and is know to originate from surface reconstructions doing the oscillations\cite{Ertl2008}. This mechanism is however not relevant at atmospheric pressures (REF?) and thus another physical mechanism must be governing the oscillations at high pressures. It has previously been shown already in 1982\cite{SALES1982} that CO oxidation at around 1\,bar will show oscillatory behavior on single crystals off Pt, Pd and Ir. Later studies has come a long way to explain the phenomenen on palladium \cite{Hendriksen2010} attributed to a cycle of surface oxide formation, oxide roughening and clean metal surface. CO oxidation oscillations has also been observed on Pt on thin-films \cite{Lund2000} and in macroscopic flow reactors \cite{Singh2010} but never on size selected platinum nanoparticles at atmospheric pressures for prolonged periods of time.

@@

\subsection{Platinum nanoparticle deposition}
Platinum nanoparticles were deposited in the reactor using a gas-aggregation source (Mantis Deposition Ltd., Nanogen 50). Platinum clusters were formed by gas-phase condensation of platinum atoms produced by impinging argon ions on a 99.99\,\% platinum target in a magnetron sputter source. After condensation the ionised fraction (60-80\%) of the clusters were size-selected by a quadropole mass selection filter according to their mass-to-charge ratio. Using this setup platinum nanoparticles with diameters in the range of 2–16 nm can be produced \cite{Nielsen2010,Nielsen2009}. The size-selected nanoparticles were after size-selection soft-landed in the reactor volume of the microreactor. The coverage was kept at approximately 0.1\% geometric coverage determined by measuring the current on the reactor during deposition. After deposition the reactors were anodicly cold-bonded \cite{Vesborg2010} to a pyrex lid to avoid sintering of the nanoparticles while bonding. Doing the experiments, samples from 3\,nm to 9\,nm has been measured, all showing oscillations.

@@

The oscillations are qualitatively similar but the oscillation period varies from a few seconds (the time-constant of the reactor is $\sim7\,$s.) up to more than an hour. An example of a slow oscillation is seen in Figure~\ref{fgr:full_oscillation}. After a short on-cycle where all CO is converted to CO$_2$ the reaction shows a fast deactivation. After the deactivation the sample quickly regains $\sim$30\% conversion. Hereafter the conversion increases linearly with time until approximately 50\% conversion is reached. Shortly after reaching 50\% conversion the sample again ignites and covert all CO to CO$_2$. After a short full conversion period the sample again deactivates and the cycle repeats itself. The ratio between periods of high and low conversion are very different among samples, but for a given sample the ratio normally does not change even though the oscillation period consistently becomes longer and longer as the experiment progresses.
\begin{figure*}
  \centering
  \includegraphics[width=17cm]{single_full_oscillation.png}
  \caption{A full oscillation period first showing a steep ignition of the sample followed by an almost immediate de-activation of the sample. For the next 65 minutes the sample slowly recovers activity until full conversion is reached again and the cycle repeats itself. This extremely long oscillation period is not commonly seen and was a result of carefull parameter tuning. Typical oscillation periods are normally between $\sim$30\,s and $\sim$30\,minutes.}
  \label{fgr:full_oscillation}
\end{figure*}

In the setup we are able to change the the pressure in the reactor volume to observe any change in behavior of the oscillations as a function of pressure. In the pressure range of 0.1\,bar to 1\,bar no change that could be attributed to the pressure was found. The oscillations does not change frequency nor qualitative behavior significantly from the oscillation shown in Figure \ref{fgr:full_oscillation}. Similarly, only a weak dependence on the CO/O$_2$ pressure was observed. At a CO/O$_2$ ratio of $\sim0.025$ only very small amplitude oscillations were observed. Increasing the CO/O$_2$ to $\sim0.175$ showed clear oscillations with a significant increase in oscillation amplitude compared to lower CO concentrations. Generally, a tendency towards slower oscillation periods for higher CO/O$_2$ ratio was observed.

@@

\section{Discussion}
The CO/O$_2$ dependence showed oscillations from a ratio of 0.025 to 0.175. Richer CO mixtures than 0.175 did not produce any oscillations in our system. CO/O$_2$ ratios lower than 0.175 were needed to provoke oscillations in our system. This ratio agrees well with literature \cite{Singh2010,Hendriksen2005} where oscillations have been observed in oxygen rich CO/O$_2$ mixtures. The low concentration of CO in the inlet gases can be attributed to the much higher sticking coefficient of CO on platinum which poisons the nanoparticle surface during the low reactivity region of the cycle. Once poisoned, an increase in temperature would be necessary to ignite the reaction again.

@@

Due to the long timescales between full conversion and low conversion plateaus the oscillations cannot be attributed to the system setup which due to its small size has timescales on the order of seconds. It is hence concluded that the oscillations is a feature related to the catalyst. It has already been shown using STM \cite{Hendriksen2002} and FT-IR \cite{Carlsson2006} as well as theoretical studies \cite{Gong2004} that the activity of Pt towards CO oxidation at atmospheric pressures is highly dependent on the oxidation state of the surface. A detailed propsal for the reaction mechanism on Pd has been developed\cite{Hendriksen2010} and our data is to a very large degree consistent with this model. According to the model, the bare metal surface is less active than the oxidized surface, and further it is proposed that a high concentration of steps will thermodynamically favor the reduced surface and a low concentration of steps will favor the oxidized surface. The oscillating behaviour now arises because the reaction on the oxidized surface will increase surface roughness and the reaction on the reduced surface will decrease surface roughness. Thus the reaction will only run a given time in the high-rate state before the catalyst will expose the reduced surface and thus lower the rate. After some time the roughness will again be sufficiently low that the catalyst will again oxidize and increase the rate. In this study, we measure on Pt nanoparticles which compared to a single crystal or a thin film, is a very rough surface. Thus it is to be expected that the particles will initially favor the reduced state, however as the reaction runs, the conditions toward the outlet of the reactor will become more reducing and at the same time, the reaction itself will smoothen the particles, further pushing towards an oxidation of the particles. Gradually as the particles towards the outlet oxidizes the turn-over will increasing and the general CO concentration will go down, further promoting more and more particles to oxidize which will gradually be seen in the mass-spectrometry as an increase in CO$_2$ and a decrase in CO. At a given point the CO concentration will become low enough that a ligh-off like phenomeneon will set in and the sample will show full conversion. In this state, most of the particles will be in the oxidzed high-activity state. As all particles are now oxidized, the will all become more and more roughened and gradually an increasing amount will go back to the reduced state. Thus less and less oxidized particles will be responsible for most of the activity and thus the roughness will increase faster and faster for these remaing active particles. Just before the overall activity drops, one must expect that only a very small fraction of the particles actually participates in the reaction and therefore the de-activation will happen very suddenly once these few particles eventually becomes sufficiently rough, that they will reduce and lose activity.

@@

>>>>

CO oxidation is one of the most studied reactions due to its importance in both automotive catalysis and in the removal of trace CO in hydrogen streams. A large effort and a large number of experiments with a variety of surface sensitive techniques have been conducted on single crystal surfaces greatly improving the understanding of this reaction. Single crystals surfaces, however, can be very much different from industrial catalysts where supported nanoparticles are used to both increase surface area hence bringing material costs down and to increase the reactivity of the material by increasing the number of active sites.

Here we report oscillatory behavior of CO oxidation on size-selected platinum nanoparticles supported on a silicon dioxide substrate at atmospheric pressure. Oscilating reactions on Platiunum at UHV-conditions has been studied extensively and is know to originate from surface reconstructions doing the oscillations\cite{Ertl2008}. This mechanism is however not relevant at atmospheric pressures and thus another physical mechanism must be governing the oscillations at high pressures. It has previously been shown already in 1982\cite{SALES1982} that CO oxidation at around 1\,bar will show oscillatory behavior on single crystals of Pt, Pd and Ir. Later studies has come a long way to explain the phenomenen on palladium \cite{Hendriksen2010} attributed to a cycle of surface oxide formation, oxide roughening and clean metal surface. CO oxidation oscillations has also been observed on Pt on thin-films \cite{Lund2000} and in macroscopic flow reactors \cite{Singh2010} but never on size selected platinum nanoparticles at atmospheric pressures for prolonged periods of time.

@@

\subsection{Platinum nanoparticle deposition}
Platinum nanoparticles were deposited in the reactor using a gas-aggregation source (Mantis Deposition Ltd., Nanogen 50). Platinum clusters were formed by gas-phase condensation of platinum atoms produced by impinging argon ions on a 99.99\% platinum target in a magnetron sputter source. After condensation the ionised fraction (60-80\%) of the clusters were size-selected by a quadropole mass selection filter according to their mass-to-charge ratio. Using this setup platinum nanoparticles with diameters in the range of 2–16 nm can be produced \cite{Nielsen2010,Nielsen2009}. The size-selected nanoparticles were after size-selection soft-landed in the reactor volume of the microreactor. The coverage was kept at approximately 0.1\% geometric coverage determined by measuring the current on the reactor during deposition. After deposition the reactors were anodicly cold-bonded \cite{Vesborg2010} to a pyrex lid to avoid sintering of the nanoparticles while bonding. Doing the experiments, samples from 3\,nm to 9\,nm has been measured, all showing oscillations.

@@

The oscillations on all samples are qualitatively similar but the oscillation period varies from a few seconds (the time-constant of the reactor is $\sim7\,$s.) up to more than an hour. An example of a slow oscillation is seen in Figure~\ref{fgr:full_oscillation}. After a short on-cycle where all CO is converted to CO$_2$ the reaction shows a fast deactivation. After the deactivation the sample quickly regains $\sim$30\% conversion. Hereafter the conversion increases linearly with time until approximately 50\% conversion is reached. Shortly after reaching 50\% conversion the sample again ignites and covert all CO to CO$_2$. After a short full conversion period the sample again deactivates and the cycle repeats itself. The ratio between periods of high and low conversion are very different among samples, but for a given sample the ratio normally does not change even though the oscillation period consistently becomes longer and longer as the experiment progresses.
\begin{figure*}
  \centering
  \includegraphics[width=17cm]{single_full_oscillation.png}
  \caption{A full oscillation period first showing a steep ignition of the sample followed by an almost immediate de-activation of the sample. For the next 65 minutes the sample slowly recovers activity until full conversion is reached again and the cycle repeats itself. This extremely long oscillation period is not commonly seen and was a result of carefull parameter tuning. Typical oscillation periods are normally between $\sim$30\,s and $\sim$30\,minutes.}
  \label{fgr:full_oscillation}
\end{figure*}

In the setup we are able to change the the pressure in the reactor volume to observe any change in behavior of the oscillations as a function of pressure. In the pressure range of 0.1\,bar to 1\,bar no change that could be attributed to the pressure was found. The oscillations does not change frequency nor qualitative behavior whith changing pressure. Similarly, only a weak dependence on the CO/O$_2$ pressure was observed. At a CO/O$_2$ ratio of $\sim0.025$ only very small amplitude oscillations were observed. Increasing the CO/O$_2$ to $\sim0.175$ showed clear oscillations with a significant increase in oscillation amplitude compared to lower CO concentrations. Generally, a tendency towards slower oscillation periods for higher CO/O$_2$ ratio was observed.

@@

\section{Discussion}
The CO/O$_2$ dependence showed oscillations from a ratio of 0.025 to 0.175. Richer CO mixtures than 0.175 did not produce any oscillations in our system. This ratio agrees well with literature \cite{Singh2010,Hendriksen2005} where oscillations have been observed in oxygen rich CO/O$_2$ mixtures. The low concentration of CO in the inlet gases can be attributed to the much higher sticking coefficient of CO on platinum which poisons the nanoparticle surface during the low reactivity region of the cycle. Once poisoned, an increase in temperature would be necessary to ignite the reaction again.

@@

Due to the long timescales between full conversion and low conversion plateaus the oscillations cannot be attributed to the reactor setup itself which due to its small size has timescales on the order of seconds. It is hence concluded that the oscillations is a feature related to the catalyst. It has already been shown using STM \cite{Hendriksen2002} and FT-IR \cite{Carlsson2006} as well as theoretical studies \cite{Gong2004} that the activity of Pt towards CO oxidation at atmospheric pressures is highly dependent on the oxidation state of the surface. A detailed propsal for the reaction mechanism on Pd has been developed\cite{Hendriksen2010} and our data is to a very large degree consistent with this model. According to the model, the bare metal surface is less active than the oxidized surface, and further it is proposed that a high concentration of steps will thermodynamically favor the reduced surface and a low concentration of steps will favor the oxidized surface. The oscillating behaviour now arises because the reaction on the oxidized surface will increase surface roughness and the reaction on the reduced surface will decrease surface roughness. Thus the reaction will only run a given time in the high-rate state before the catalyst will expose the reduced surface and thus lower the rate. After some time the roughness will again be sufficiently low that the catalyst will again oxidize and increase the rate. In this study, we measure on Pt nanoparticles which compared to a single crystal or a thin film, is a very rough surface. Thus it is to be expected that the particles will initially favor the reduced state, however as the reaction runs, the conditions toward the outlet of the reactor will become more reducing and at the same time, the reaction itself will smoothen the particles, further pushing towards an oxidation of the particles. Gradually as the particles towards the outlet oxidizes the turn-over will increasing and the general CO concentration will go down, further promoting more and more particles to oxidize which will gradually be seen in the mass-spectrometry as an increase in CO$_2$ and a decrase in CO. At a given point the CO concentration will become low enough that a ligh-off like phenomeneon will set in and the sample will show full conversion, if not all particles where oxidized at this point, they will oxidize much faster after light-off due to the much lower CO-concentration, the increased rate will not be visible in the mass-spectromery since reactor is allready in full conversion. As most of the particles are now oxidized, the will all become more and more roughened and gradually an increasing amount will go back to the reduced state. Thus less and less oxidized particles will be responsible for most of the activity and thus the roughness will increase faster and faster for these remaing active particles. Just before the overall activity drops, one must expect that only a very small fraction of the particles actually participates in the reaction and therefore the de-activation will happen very suddenly once these few particles eventually becomes sufficiently rough, that they will reduce and lose activity.

@@

\section{Conclusion}
We have for the first time demonstrated oscillationg CO oxidation on Pt nanoparticles, measured with high-quality mass-spectrometry. We have shown that the phenomenon is happining for a wide range of sizes of nanoparticles and for a relatively large span i CO-concentration.
